---
title: "Airline Data analysis"
output: html_document
date: "2023-08-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
Invistico<-read.csv("Invistico_Airline.csv") #read the data set into R environment
head(Invistico) #list first 6 rows of the data set
str(Invistico) #Check the structure of the data
New_Invistico<-Invistico %>% na.omit() #omit all the missing values

```

```{r}
#Visualize the relationship between late arrival and flight distance
New_Invistico %>%ggplot(aes(x = Flight.Distance, y = Arrival.Delay.in.Minutes)) +
  geom_point(color = "blue") +
  labs(title = "Flight Distance vs. Arrival Delay",
       x = "Flight Distance",
       y = "Arrival Delay (minutes)")

cor.test(New_Invistico$Flight.Distance,New_Invistico$Arrival.Delay.in.Minutes)
```
There is no meaningful relationship between the two variables. A correlation test further reveals that the relationship between the two variables is statistically significant, at least at the 5% level of significance. Although, an increase in flight distance is associated with an increase in arrival delay, even though the relationship is weak.

```{r}
New_Invistico %>% ggplot(aes(x = satisfaction, fill = Class)) +
  geom_bar(position = "dodge") +
  labs(title = "Satisfaction by Class",
       x = "Satisfaction",
       y = "Count",
       fill = "Class")
```

Most respondents claiming to be satisfied are generally flying the business class. The persons claiming dissatisfaction are mostly in the Eco class.

```{r}
New_Invistico %>% ggplot(aes(x = satisfaction, fill = Gender)) +
  geom_bar(position = "dodge") +
  labs(title = "Satisfaction by Gender",
       x = "Satisfaction",
       y = "Count",
       fill = "Gender")
```
Females report more satisfaction compared to males and more males reporting dissatisfaction compared to females. As such, it is intuitive to explore if there is a relationship between gender and the class one flies.


```{r}
New_Invistico %>% ggplot(aes(x = Gender, fill = Class)) +
  geom_bar(position = "dodge") +
  labs(title = "Gender by Class",
       x = "Gender",
       y = "Count",
       fill = "Class")
```

Both genders tend to fly all classes at almost the same proportions. Therefore, no gender has preferences to a particular class of flights. 


```{r}
## For the following section, we attempt to develop a prediction model for satisfaction based on cleanliness, ease of online booking, food and drink, inflight entertainment, inflight wifi service, leg room, on-board service, online support and seat comfort.

predictor_vars <- c("Seat.comfort","Departure.Arrival.time.convenient","Food.and.drink","Gate.location", "Inflight.wifi.service", "Inflight.entertainment", "Online.support","Ease.of.Online.booking", "On.board.service", "Leg.room.service", "Baggage.handling","Checkin.service","Cleanliness","Online.boarding")

New_Invistico$satisfaction <- factor(New_Invistico$satisfaction,levels = c("dissatisfied", "neutral", "satisfied"), labels = c(0, 1, 2))

levels(New_Invistico$satisfaction)

#Train-Test Split
set.seed(42) # For reproducibility
train_index <- createDataPartition(New_Invistico$satisfaction, p = 0.7, list = FALSE)
train_data <- New_Invistico[train_index, ]
test_data <- New_Invistico[-train_index, ]
```

